---
title: "Homework 10"
author: "Emily Beasley"
date: "October 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(ggplot2)
```

Problems 1 and 2 are taken from Homework 9, since they're the same problems with different variable names.
```{r}
vector1 <- sample(-3:3, size = 10, replace = T)

#using a for loop
newvec <- logical()
for(i in 1:length(vector1)){
  ifelse(vector1[i] == 0, newvec[i] <- 1, newvec[i] <- 0)
}

sum(newvec)

#using a single line
length(which(vector1 == 0))
```

Problem 3. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
```{r}
rows <- 2
cols <- 3

mat <- matrix(NA, nrow = rows, ncol = cols)
for(i in 1:rows){
  for(j in 1:cols){
    mat[i,j] <- i*j
  }
}

print(mat)
```

Problem 4. Use the code from yesterday’s class to design and conduct a randomization test for some of your own data. To finish this assignment in a reasonable amount of time, I'm going to use some data from my master's thesis and see if the species richness/area slope I found is a significant departure from randomness.
```{r}
#Set the seed
set.seed(15)

#Load global variables
areas <- readRDS(file = "area.rds")
msom.out <- readRDS(file = "modelsanscovs.rds") #Note: this actually had covs

nsims <- 1000

#Get species richness counts
get.richness <- function(x){
  all.rich <- x$sims.list$Nsite
  mean.Nsite <- apply(all.rich, 2, mean)
  
  return(mean.Nsite)
}

Nsite <- get.richness(x = msom.out)

#put it all into a data.frame
rich.data <- data.frame(ID = 1:length(Nsite), areas, Nsite)

#Calculate slope of area/richness regression
getmetric <- function(z){
  . <- lm(z[,3]~z[,2])
  . <- summary(.)
  . <- .$coefficients[2,1]
  slope <- .
  
  return(slope)
}

obs.slope <- getmetric(z = rich.data)

#Shuffle the data 1000 times to get some random-ish data sets
shuffle <- function(z, nsims){
  shuffled.data <- list()
  for(i in 1:nsims){
   shuffled.data[[i]] <- data.frame(z[,1:2], sample(z[,3]))
   
  }
  return(shuffled.data)
}

shuffled.data <- shuffle(z = rich.data, nsims = nsims)

shuffled.slope <- lapply(shuffled.data, getmetric)

#Calculate p-value from simulated metrics
get.pval <- function(obs, shuffled){
  p.lower <- mean(shuffled <= obs)
  p.upper <- mean(shuffled >= obs)
  return(c(p.lower, p.upper))
}

pval <- get.pval(obs = obs.slope, shuffled = shuffled.slope)
print(pval)

#P-value is 0.011, so the observed slope is significantly higher

#Create function to set up data for plotting
get.frame <- function(x){
  sims <- unlist(x)
  df <- data.frame(ID = seq_along(sims), sims)
  
  return(df)
}

df <- get.frame(shuffled.slope)

#Now plot it
ggplot(data = df, aes(x = df$sims))+
  geom_histogram(aes(fill = I("limegreen"), color = I("black")))+
  geom_vline(aes(xintercept = obs.slope, col = "blue"))+
  theme_bw()
```

Problem 5: Compare the p-value above (0.011) with the p-value from an actual linear regression. If they're fairly different, change the starting seed or number of replications.

```{r}
linear <- lm(data = rich.data, Nsite~areas)
print(summary(linear))

#The p-value here is 0.026, which I think is close enough.
```