---
title: "Homework 4"
author: "Emily Beasley"
date: "September 19, 2018"
output: html_document
---

###Working with Vectors

1. Some expressions
```{r, echo = T, warning = F}
x <- 1.1
a <- 2.2
b <- 3.3

z <- x^(a^b)
print(z)

z <- (x^a)^(b)
print(z)

z <- 3*(x^3) + 2*(x^2) + 1
print(z)
```

2. Create some vectors
```{r, echo = T, warning = F}
a <- c(seq(1:8), seq(from = 7, to = 1))
a

b <- c(1, rep(2,2), rep(3,3), rep(4,4), rep(5,5))
b

c <- c(5, rep(4,2), rep(3,3), rep(2,4), rep(1,5))
c
```

3. Applying functions to vectors
```{r, echo = T, warning = F}
set.seed(5)

xy<-runif(n = 2)
xy

polar <- c(sqrt((xy[1]^2)+(xy[2]^2)), atan(xy[2]/xy[1]))
polar
```

4. Updating vectors
```{r, echo = T, warning = F}
queue <- c("sheep", "fox", "owl", "ant")

queue[5] <- "serpent"
queue

queue <- queue[-1]
queue

queue <- c("donkey", queue)
queue

queue <- queue[-5]
queue

queue <- queue[-3]
queue

queue <- c(queue[1:2], "aphid", queue[3])
queue

which(queue == "aphid")
```

5. Creating more complex sequences
```{r, echo = T, warning = F}

asequence <- seq(1:100)

nope <- c(seq(from = 2, to = 100, by = 2), seq(from = 3, to = 100, by = 3),
          seq(from = 7, to = 100, by = 7))
nope <- unique(nope)

my.seq <- setdiff(asequence, nope)
my.seq
```

###Regular expressions

6. Change this:
```text
First String    Second      1.22      3.4
Second          More Text   1.555555  2.2220
Third           x           3         124
```

Into this:
```text
First String,Second,1.22,3.4
Second,More Text,1.555555,2.2220
Third,x,3,124
```

By using this:
```text
find: (\S+?(?=\s{1, }))\s{1, }
replace with: \1,
```
Here, I searched for 1 or more consecutive non-space characters `/S+`, stopping when it found a string of more than one space character `?(?=\s{1, })`. I added the sequence of multiple spaces at the end of the search string `\s{1, }`. Then I changed the strings of multuple space characters to commas.

7. Change this:
```text
Ballif, Bryan, University of Vermont
Ellison, Aaron, Harvard Forest
Record, Sydne, Bryn Mawr
```
To this:
```text
Bryan Ballif (University of Vermont)
Aaron Ellison (Harvard Forest)
Sydne Record (Bryn Mawr)
```
I used this:
```text
find: (\w*),\s(\w*),\s(.*)
replace with: \2 \1 \(\3\)
```
Capture string of 1 or more word characters, then a comma, then a space, then one or more word characters, then space, then the rest of the line. Replace with second, then first capture, separated by spaces. To include parentheses, use "\)" to open, then "\)" to close.

8. Change this:
```text
0001 Georgia Horseshoe.mp3 0002 Billy In The Lowground.mp3 0003 Cherokee Shuffle.mp3 0004 Walking Cane.mp3
```

To this:
```text
Georgia Horseshoe_0001.mp3
Billy In The Lowground_0002.mp3
Cherokee Shuffle_0003.mp3
Walking Cane_0004.mp3
```
Using this:
```text
find: (\d+)\s(.*?(?=[.mp3]))(.mp3)\s
replace: \2_\1\3\n
```
I searched for and chaptured a number string of 1 or more characters `(\d+)`, followed by a non-captured space, then captured a string of any character until it hit ".mp3" `(.*?(?=[.mp3]))`. I captured the ".mp3" after the character string, and included the space separating tracks in the search. For replacement, I separated the second and first catch with an underscore, and added a line break `\n` at the end.

9. Change this:
```text
Camponotus,pennsylvanicus,10.2,44
Camponotus,herculeanus,10.5,3
Myrmica,punctiventris,12.2,4
Lasius,neoniger,3.3,55
```
To this:
```text
C_pennsylvanicus,44
C_herculeanus,3
M_punctiventris,4
L_neoniger,55
```
Using this:
```text
find: (\w)\w+,(\w+),(\d+\.\d+),(\d+)
replace: \1_\2,\4
```
I searched for a single word character followed by 1 or more word characters, then a comma, then 1 or more word characters. After a second comma I searched for two groups of numeric characters separated by a decimal. After a third comma I searched for one or more numeric characters.

Then change the original text to this:
```text
C_penn,44
C_herc,3
M_punc,4
L_neon,55
```

Using this:
```text
find: (\w)\w+,(\w{4})\w+,(\d+\.\d),(\d+)
replace: \1_\2,\4
```

Same as above, except I changed the second capture to only include the first 4 letters of the word string. I searched for the rest of the letters in the second word string, but left them out of the capture.