---
title: "Homework 6"
author: "Emily Beasley"
date: "October 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
library(ggplot2)
library(MASS)
library(tidyverse)
```
Code for generated data is comment out. All outputs are of data from my master's thesis.

Generate or read in data:
```{r}
# #Generate a truncated normal distribution
# z <- rnorm(n=3000,mean=0.2)
# z <- data.frame(1:3000,z)
# names(z) <- list("ID","myVar")
# z <- z[z$myVar>0,] #This line truncates the distribution so all values > 0
# str(z)
# summary(z$myVar)

mamm <- read.csv(file = "C:/users/beasley/dropbox/Thesis Stuff/summer mamm.csv",
                   sep = ",")

mamm %>%
  filter(Species == "Sigmodon hispidus", Mass != ">300"|Mass != "") %>%
  select(Mass) %>%
  droplevels() %>%
  {. ->> fatrat}

fatrat <- data.frame(1:length(fatrat), fatrat)
fatrat$Mass <- as.numeric(as.character(fatrat$Mass))
str(fatrat)
summary(fatrat$Mass)
```

Plot a histogram of the data and add density curve
```{r}
# plot.gen <- ggplot(data=z, aes(x=myVar, y=..density..)) +
#   geom_histogram(color="grey60",fill="cornsilk",size=0.2)+
#   geom_density(linetype="dotted",size=0.75)
# 
# plot.gen

ratplot <- ggplot(data = fatrat, aes(x = Mass, y = ..density..))+
  geom_histogram(color = "grey60", fill = "grey", size = 0.2)+
  geom_density(linetype = "dotted", size = 0.75)+
  theme_bw()

ratplot  

#It's bimodal. Huzzah.
```

Get maximum likelihood parameters for a normal distribution:
```{r}
#Fit the distribution
# normPars <- fitdistr(z$myVar,"normal")
# #Look at it
# print(normPars)
# str(normPars)
# normPars$estimate["mean"]

normPars <- fitdistr(fatrat$Mass,"normal")
#Look at it
print(normPars)
str(normPars)
normPars$estimate["mean"]
```

Plot probability density
```{r}
#Pull mean and stdev from data
# meanML <- normPars$estimate["mean"]
# sdML <- normPars$estimate["sd"]
# 
# #Generate x values
# xval <- seq(0,max(z$myVar),len=length(z$myVar))
# 
# #This fits a curve based on the distribution function ("fun") and the 
# #parameters you provide ("args")
# stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", 
#                        n = length(z$myVar), args = list(mean = meanML, sd = sdML))
# plot.gen + stat

meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

#Generate x values
xval <- seq(0,max(fatrat$Mass),len=length(fatrat$Mass))

#This fits a curve based on the distribution function ("fun") and the 
#parameters you provide ("args")
stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", 
                       n = length(fatrat$Mass), args = list(mean = meanML, 
                                                            sd = sdML))
ratplot + stat
```

Plot more probability densities
```{r}
# #Exponential
# expoPars <- fitdistr(z$myVar,"exponential")
# rateML <- expoPars$estimate["rate"]
# 
# stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", 
#                        n = length(z$myVar), args = list(rate=rateML))
# plot.gen + stat + stat2
# 
# #Uniform
# stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", 
#                        n = length(z$myVar), args = list(min=min(z$myVar), 
#                                                         max=max(z$myVar)))
# plot.gen + stat + stat2 + stat3
# 
# #Gamma
# gammaPars <- fitdistr(z$myVar,"gamma")
# shapeML <- gammaPars$estimate["shape"]
# rateML <- gammaPars$estimate["rate"]
# 
# stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", 
#                        n = length(z$myVar), args = list(shape=shapeML,
#                                                         rate=rateML))
# plot.gen + stat + stat2 + stat3 + stat4
# 
# #Beta
# plotgen.Special <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), 
#                                       y=..density..)) +
#   geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
#   xlim(c(0,1)) +
#   geom_density(size=0.75,linetype="dotted")
# 
# betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),
#                      start=list(shape1=1,shape2=2),"beta")
# shape1ML <- betaPars$estimate["shape1"]
# shape2ML <- betaPars$estimate["shape2"]
# 
# statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, 
#                              colour="orchid", n = length(z$myVar), 
#                              args = list(shape1=shape1ML,shape2=shape2ML))
# plotgen.Special + statSpecial
# 
# #Gamma and beta look pretty good

#Exponential
expoPars <- fitdistr(fatrat$Mass,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", 
                       n = length(fatrat$Mass), args = list(rate=rateML))
ratplot + stat + stat2

#Uniform
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", 
                       n = length(fatrat$Mass), args = list(min=min(fatrat$Mass), 
                                                        max=max(fatrat$Mass)))
ratplot + stat + stat2 + stat3

#Gamma
gammaPars <- fitdistr(fatrat$Mass,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", 
                       n = length(fatrat$Mass), args = list(shape=shapeML,
                                                        rate=rateML))
ratplot + stat + stat2 + stat3 + stat4

#Beta
ratplot.Special <- ggplot(data=fatrat, aes(x=Mass/(max(Mass + 0.1)), 
                                      y=..density..)) +
  geom_histogram(color="grey60",fill="grey",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")+
  theme_bw()

betaPars <- fitdistr(x=fatrat$Mass/max(fatrat$Mass + 0.1),
                     start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, 
                             colour="orchid", n = length(fatrat$Mass), 
                             args = list(shape1=shape1ML,shape2=shape2ML))
ratplot.Special + statSpecial

#And nothing fits because it's bimodal
```

Maximum likelihood of beta distribution
```{r}
betaPars <- fitdistr(x=fatrat$Mass/max(fatrat$Mass + 0.1),
                     start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

#Generate beta distribution with the above parameters
simdata <- rbeta(n = length(fatrat$Mass), shape1 = shape1ML, shape2 = shape2ML)

simframe <- data.frame(1:length(simdata), simdata)

#Plot the simulated data
ggplot(data = simframe, aes(x = simdata, y = ..density..))+
  geom_histogram(color = "grey60", fill = "grey", size = 0.2)+
  geom_density(linetype = "dotted", size = 0.75)+
  theme_bw()

ratplot
```

The two plots (simulated & real data) are wildly different due to the bimodal distribution of the real data. This data set includes both juvenile/subadult and adult cotton rats, which may be the cause of the two peaks: one for juvenile mass, the other for adult mass. Choosing one or the other may yield a data set that fits a unimodal distribution fairly well.