---
title: "Homework 7"
author: "Emily Beasley"
date: "October 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
library(msm)
library(tidyverse)
```

The following data sim is a little different from the instructions on the course page. I wanted to do a simulation that was more applicable to my dissertation research. I will simulate some occupancy data for ectoparasites on mammalian hosts that will eventually be used to build a multi-scale occupancy model. Since the model will take a long time to figure out, for this assignment I'm going to demonstrate how imperfect detection can affect ectoparasite richness counts.

First, generate some preliminary variables.
```{r}
set.seed(15)

#Mammal species
mammspecs <- data.frame(as.character(seq(1:10)))
colnames(mammspecs) <- "mamm.specs"

#assign fake genera to mammals. This will be important later,
#ectos tend to be phylogenetically conserved
mammspecs$genus <- c(rep("genus1", 5), rep("genus2", 5))

#Ecto species
ectospecs <- data.frame(as.character(seq(1:12)))
colnames(ectospecs) <- "ecto.specs"
#Twelve ectoparasite species is a low-ish number, but it'll do

#Assign ectos as generalist, specialist, or neither
ectospecs$life.history <- c(rep("generalist", 4), rep("specialist", 4), 
                            rep("neither", 4))
```

Generate occupancy probabilities using this nasty function.
```{r}
makin.data <- function(){
  #Create empty objects
  genus <- logical()
  occprobs <- matrix(NA, nrow = 12, ncol = 10)
  
  for(i in 1:length(ectospecs$ecto.specs)){
  #Assign a mammal genus to the ecto species
  genus[i] <- sample(mammspecs$genus, size = 1)
  
  #Generalist species: can occupy all species in a genus
  if(ectospecs$life.history[i] == "generalist"){
    #Generate occupancy probs for all mamm in the genus
    for(j in 1:length(mammspecs$mamm.specs)){
      if(mammspecs$genus[j] == genus[i]){
        occprobs[i,j] <- runif(1, 0, 1)
      } else{
        occprobs[i,j] <- 0
      }
    }
    
  #Specialist species: Only one host  
  } else if(ectospecs$life.history[i] == "specialist"){
    #Generate an occupancy prob for a single host
    host <- sample(mammspecs$mamm.specs, size = 1)
    for(j in 1:length(mammspecs$mamm.specs))
      if(mammspecs$mamm.specs[j] == host){
        occprobs[i,j] <- runif(1,0,1)
      } else{
        occprobs[i,j] <- 0
      }
  
  #"Neither" category: one primary and one secondary host
  } else if(ectospecs$life.history[i] == "neither"){
      #Generate occupancy prob for a primary and secondary host
        sample <- sample(which(mammspecs$genus == genus[i]), 2)
        occprobs[i, sample[1]] <- runif(1, 0.5, 1) #primary host
        occprobs[i, sample[2]] <- runif(1, 0, 0.5) #secondary host
        occprobs[is.na(occprobs)] <- 0
  }
  }
  return(occprobs)
}

some.data <- makin.data()
```

Now I have occupancy data for each host species. The problem: for the multiscale model to work, I need to treat each host population as a site. So I'll draw a probability for each population using the host species probability as a mean.
```{r}
#Decide how many physical sites to "sample" from
n.site <- 20 #This is ususally the minimum for an effective abundance model

#Create a vector with the name of each host population
write.names <- function(x){
  spec.names <- logical(length = length(x))
  for(i in 1:length(x)){
    spec.names[i] <- as.character(paste("Spec",x[i], sep = "")) 
  }
  return(spec.names)
} 

write.sites <- function(x){
  site.names <- logical(length = x)
  for(i in 1:x){
    site.names[i] <- as.character(paste("Site",i,sep = ""))
  }
  return(site.names)
}

names <- write.names(x = mammspecs$mamm.specs)
sites <- write.sites(x = n.site)

#Generate an occupancy prob for each host population based on 
#the species-level occupancy prob
spec.probs.array <- function(x,y,z){
pop.array <- array(dim = c(length(x), y,length(z)))

for(i in 1:length(x)){
  for(j in 1:y){
    for(k in 1:length(z)){
      if(some.data[k,i] != 0){
          pop.array[i,j,k] <- rtnorm(n = 1, mean = some.data[k,i], sd = 0.05, 
                                    lower = 0, upper = 1)
        #Use a truncated normal distribution to make sure it's a probability
        #May use a different distribution later
      } else{
        pop.array[i,j,k] <- 0
    }
  }
  }
}
return(pop.array)
}

poparray <- spec.probs.array(x = mammspecs$mamm.specs, y = n.site, 
                            z = ectospecs$ecto.specs)

#Function to convert array into data frame
spec.probs.mat <- function(x){
  popoccs <- matrix(NA, nrow = dim(x)[1]*dim(x)[2], ncol = dim(x)[3])
  for(i in 1:dim(x)[3]){
    popoccs[,i] <- as.vector(x[,,i])
  }
  popoccs <- as.data.frame(popoccs)
  return(popoccs)
}

popmat <- spec.probs.mat(x = poparray)
```

Use the occupancy probabilities to generate sample presence-absence data.
```{r}
get.occs <- function(x){
  occstate <- matrix(NA, nrow = length(x[,1]), ncol = length(x[1,]))
  for(i in 1:length(x[,1])){
    for(j in 1:length(x[1,])){
      occstate[i,j] <- rbernoulli(n = 1, p = popmat[i,j])
      occstate[which(occstate[i,j] == TRUE)] <- 1
      occstate[which(occstate[i,j] == FALSE)] <- 0
    }
  }
  return(occstate)
}
  
occstate <- get.occs(x = popmat)

#Now combine all of this nonsense into a single data frame
specnames <- rep(names, length(sites))

sitenames <- list()
for(i in 1:length(sites)){
  sitenames[[i]] <- rep(sites[i], length(names))
}
sitenames <- unlist(sitenames)

occ.mat <- cbind(names, sitenames, occstate)

#Last thing: convert matrix to long form
occ.mat %>%
  as.data.frame() %>%
  gather(V3:V14, key = "Ecto", value = "Prob") %>%
  {. ->> long.occ}
```


```{r}

```
