---
title: "Homework 8"
author: "Emily Beasley"
date: "October 31, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
set.seed(15)
library(msm)
library(tidyverse)
```

The following code condenses the simulation from Homework 7. I will use it to determine the minimum number of sites needed to detect all 12 ectoparasite species.

Start by building global variables:
```{r Global Variables}
n.mamm <- 10
mamm.genera <- c("Genus1", "Genus2")
mamm <- readRDS(file = "truemamm.rds")

n.ecto <- 12
life.history <- c("generalist", "specialist", "neither")

n.site <- 1:30
```

Generate occupancy probabilities for each host/ecto pair.
```{r Host Species Probs}
spec.occ.probs <- function(n.mamm, genera, n.ecto, life.history){
  #Generate name vectors for mammals and ectos
  mammspec <- logical()
  for(i in 1:n.mamm){
    mammspec[i] <- as.character(paste("Spec", i, sep = ""))
  }
  
  ectospec <- logical()
  for(i in 1:n.ecto){
    ectospec[i] <- as.character(paste("Spec", i, sep = ""))
  }
  
  #Generate mammal and ecto data frames
  mammdat <- data.frame(mammspec = mammspec,
                        mammgenus <- sample(genera, size = n.mamm, replace = T))
  ectodat <- data.frame(ectospec = ectospec, 
                        ectolife = sample(life.history, size = n.ecto, replace = T))
  
  #Generate occupancy probabilities. Commence for loop of hell.
  genus <- logical()
  occprobs <- matrix(NA, nrow = n.ecto, ncol = n.mamm)
  
  for(i in 1:n.ecto){
    #Assign a mammal genus to the ecto species
    genus <- sample(mammdat$mammgenus, size = 1)
  
    #Generalist species: can occupy all species in a genus
    if(ectodat$ectolife[i] == "generalist"){
      #Generate occupancy probs for all mamm in the genus
      for(j in 1:n.mamm){
        if(mammdat$mammgenus[j] == genus){
          occprobs[i,j] <- runif(1, 0, 1)
        } else{
          occprobs[i,j] <- 0
        }
      }
    
    #Specialist species: Only one host  
    } else if(ectodat$ectolife[i] == "specialist"){
      #Generate an occupancy prob for a single host
      host <- sample(mammdat$mammspec, size = 1)
      for(j in 1:length(mammdat$mammspec))
        if(mammdat$mammspec[j] == host){
          occprobs[i,j] <- runif(1,0,1)
        } else{
          occprobs[i,j] <- 0
        }
  
    #"Neither" category: one primary and one secondary host
    } else if(ectodat$ectolife[i] == "neither"){
        #Generate occupancy prob for a primary and secondary host
        somehosts <- sample(which(mammdat$mammgenus == genus), size = 2)
        occprobs[i, somehosts[1]] <- runif(1, 0.5, 1) #primary host
        occprobs[i, somehosts[2]] <- runif(1, 0, 0.5) #secondary host
        occprobs[is.na(occprobs)] <- 0
  }
  }
  return(occprobs)
}

host.probs <- spec.occ.probs(n.mamm, mamm.genera, n.ecto, life.history)
```

Generate occupancy probabilities for each ecto on each host population. Probs are drawn from a normal distribution with the host species/ecto species probability as the mean.
```{r Host Pop Probs}
pop.occ.probs <- function(x = n.mamm,y,z = n.ecto,mat){
  pop.array <- array(dim = c(x, y, z))

  for(i in 1:x){
    for(j in 1:y){
      for(k in 1:z){
        if(mat[k,i] != 0){
            pop.array[i,j,k] <- rtnorm(n = 1, mean = mat[k,i], sd = 0.05, 
                                    lower = 0, upper = 1)
          #Use a truncated normal distribution to make sure it's a probability
          #May use a different distribution later
        } else{
          pop.array[i,j,k] <- 0
      }
    }
    }
  }
  
  #Convert array to a matrix
  popoccs <- matrix(NA, nrow = x*y, ncol = z)
  for(i in 1:z){
    popoccs[,i] <- as.vector(pop.array[,,i])
  }
  popoccs <- as.data.frame(popoccs)
}

pop.probs <- pop.occ.probs(y = n.site[4], mat = host.probs)
```

Use the occupancy probabilities to get presence-absence data
```{r Occupancy State}
get.occs <- function(x = n.mamm,y,z = n.ecto,mat){
  occstate <- matrix(NA, nrow = x*y, ncol = z)
  for(i in 1:(x*y)){
    for(j in 1:z){
      occstate[i,j] <- as.numeric(rbernoulli(n = 1, p = mat[i,j]))
    }
  }
  return(occstate)
}

occstate <- get.occs(y = n.site[4], mat = pop.probs)
```

Simulate the detection process, where individual hosts are replicates of the population.
```{r Detection Data}
det.data <- function(mamms, n.site, n.ecto, mat){
  #Sample from a smaller number of sites
  mamm.sites <- sample(seq(length(mamms[1,])), size = n.site, replace = F)
  mammZ <- mamms[,mamm.sites]

  #Simulate mean detection prob for each ectoparasite
  #Same det prob regardless of host species
  ecto.dets <- runif(n = n.ecto, min = 0, max = 0.2)
  #Detection probs for ectoparasites in reality are usually low
  #Parasites also tend to aggregate, but for now I'm not messing with that

  #Create vector with numbers of sampling events K
  Ks <- as.vector(mammZ)
  maxK <- max(Ks)

  #Create detection histories
  L <- list()

  for(i in 1:n.ecto){
    y <- matrix(NA, nrow = length(Ks), ncol = maxK)
    for(k in 1:length(Ks)){
      for(j in 1:Ks[k]){
        y[k,j] <- as.numeric(rbernoulli(n = 1, p = ecto.dets[i]*mat[k,i]))
      }
    }
    L[[i]] <- y
  }

  #Change list of matrices into array
  obsdata<-array(unlist(L), dim=c(length(Ks), maxK, n.ecto))
  #Change NA to 0
  obsdata[is.na(obsdata)] <- 0

  #Add all observed occurances; if there is more than 1, change to 1
  obsdata2 <- apply(obsdata, c(1,3), sum)
  obsdata2[obsdata2 >= 1] <- 1

  return(obsdata2)
}
```

Get regional ectoparasite richness.
```{r Det Richness}
det.rich <- function(mat){
  . <- colSums(mat)
  reg.rich <- length(which(. > 0))
}
```

Putting it all together: How many sites do I need to sample to detect all 12 ectoparasite species?
```{r}
ectorich <- logical()
for(i in seq_along(n.site)){
  . <- spec.occ.probs(n.mamm = n.mamm, genera = mamm.genera, n.ecto = n.ecto, 
                      life.history = life.history)
  . <- pop.occ.probs(y = n.site[i], mat = .)
  . <- get.occs(y = n.site[i], mat = .)    
  . <- det.data(mamms = mamm, n.ecto = n.ecto, n.site = n.site[i], mat = .)
  . <- det.rich(mat = .)
  ectorich[i] <- .
}

print(which(ectorich == 12))
```

Based on this sim, it would take at least 20 sites to consistently detect all ectoparasite species. Even then, it doesn't always happen.

Or you could data-augment an MSOM to model undetected species.